---
title: "replicate PPH paper"
output: pdf_document
---

```{r}
library(tidyverse)
library(tidymodels)
library(readr)
PPH <- read_csv("PPH_161cols.csv")

race_table <- data.frame (race = c(1,2,3,4,5,6,NA),
                  race_description = c("White", "Black", "Hispanic", "Asian or PI", "Native American", "Other", "Unknown") )

PPH <- left_join(PPH, race_table, by = "race" )

head(PPH)
```

"A total of 360,370 (3.2%) of 11,260,869
estimated deliveries to women aged
15e54 years complicated by postpartum
hemorrhage were included in this anal-
ysis (Table 1)."

```{r}
# We have 117583 observation of PPH patients.
# If we take weight into consideration, then we have

PPH %>% select(discwt) %>% sum()
```
PPH patient count by racce
```{r}

race_stats <- PPH %>% group_by(race_description) %>% summarise(count = n(), demograph_perc = round(n()/117583, digits = 3)) %>% 
  arrange(desc(demograph_perc)) %>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"All")))

race_stats$race_description <- factor(race_stats$race_description,levels = c("Black", "White", "Hispanic", "Asian or PI", "Unknown", "Native American","Other", "All"))


```

```{r}
library(scales)
ggplot(race_stats[-c(5,8),], aes(x = race_description, y = demograph_perc)) +
  geom_bar(stat = "identity", width= 0.8) +
  xlab("Race") + 
  ylab("Percentage ") + 
  scale_fill_brewer() +
  scale_y_continuous(labels = scales::percent) + 
  theme(axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8))

```

### PPH Mortality by Race

```{r}
died <- PPH %>% filter(died == 1) %>% group_by(race_description) %>% summarise(mortality = n())%>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"All")))

died <- left_join(race_stats, died, by = "race_description")
died <- died %>% 
  mutate(mort_rate = mortality/count)

died$race_description <- factor(died$race_description,levels = c("Black", "White", "Hispanic", "Asian or PI", "Unknown", "Native American","Other", "All"))


ggplot(died[-c(5),], aes(x = race_description, y = mort_rate)) +
  geom_bar(stat = "identity", width= 0.8, fill = "blue") +
  scale_fill_brewer()+
  xlab("Race") + 
  ylab("Mortality Rate") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(axis.text.x = element_text(size = 7), 
        axis.text.y = element_text(size = 8))

```

### Fig 1.A
The risk for severe maternal morbidity (SMM) based on race, including transfusion 

```{r}

severe_21 <- PPH %>% filter(SMM == "Y") %>%
  group_by(race_description) %>% 
  summarise(severe = n())%>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"All")))

severe_21 <- left_join(died, severe_21, by = "race_description")

severe_21$race_description <- factor(severe_21$race_description,levels = c("Black", "White", "Hispanic", "Asian or PI", "Unknown", "Native American","Other", "All"))


severe_21 <- severe_21 %>% 
  mutate(smm_rate = severe/count)


ggplot(severe_21[-c(5),], aes(x = race_description, y = smm_rate)) +
  geom_bar(stat = "identity", width= 0.5, fill = "#21aca9") +
  scale_fill_gradient()+
  xlab("Race") + 
  ylab("Severe Maternal Morbidity Rate") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 0.3)) + 
  scale_x_discrete(labels=c("Black" = "Black*", "White" = "White(ref)", "Hispanic" =  "Hispanic*", "Asian or PI"= "Asian or PI*", "Native American" = "Native American**","Other" = "Other*", "All"))+
 # geom_text(aes(label = severe_21_perc),size = 3, vjust = -0.5, position = position_dodge(0.9)) +
  theme(axis.text.x = element_text(size=7,angle = 45, vjust = 1, hjust=1),
        axis.text.y = element_text(size = 8),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "grey80"),
       panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey80"),
        panel.background = element_blank())
       #panel.border = element_rect(size = 0.5, linetype = 'solid', colour = "grey80", fill = NA))

```



### Figure 1.B

above without transfusion

```{r}
severe_20 <- PPH %>% filter(SMM == "Y", cdc_s18 == 0) %>%
  group_by(race_description) %>% 
  summarise(severe_no_trans = n())%>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"All")))

severe_20 <- left_join(severe_21, severe_20, by = "race_description")

severe_20 <- severe_20 %>% 
  mutate(smm_no_trans_rate = severe_no_trans/count)

ggplot(severe_20[-c(5),], aes(x = race_description, y = smm_no_trans_rate)) +
  geom_bar(stat = "identity", width= 0.5, fill = "blue") +
  xlab("Race") + 
  ylab("Severe morbidity excluding transufusion" ) + 
  scale_y_continuous(labels = scales::percent) + 
  theme(axis.text.x = element_text(size = 7), 
        axis.text.y = element_text(size = 8),
        axis.title = element_text(size=10)
        )
```

### Fig.2
The risk for disseminated intravascular coagulation (A), transfusion (B), and hysterectomy (C) within a race for patients with postpartum hemorrhage.

```{r}
dic <- PPH %>% filter(cdc_s8 == 1) %>%
  group_by(race_description) %>% 
  summarise(dic_count = n()) %>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"All")))

dic <- left_join(severe_20, dic, by = "race_description")

dic <- dic %>% 
  mutate( dic_risk = dic_count/count)

#transfusion
trans <- PPH %>% filter(cdc_s18 == 1) %>%
  group_by(race_description) %>% 
  summarise(trans_count = n()) %>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"All")))


trans <- left_join(dic, trans, by = "race_description")

trans <- trans %>% 
  mutate( trans_risk = trans_count/count)

#hyst

hyst <- PPH %>% filter(cdc_s19 == 1) %>%
  group_by(race_description) %>% 
  summarise(hyst_count = n()) %>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"All")))

hyst <- left_join(trans, hyst, by = "race_description")

hyst <- hyst %>% 
  mutate( hyst_risk = hyst_count/count)

hyst$race_description <- factor(hyst$race_description,levels = c("Black", "White", "Hispanic", "Asian or PI", "Unknown", "Native American","Other", "All"))

stack <- hyst %>% 
  select(race_description,trans_risk,dic_risk, hyst_risk) %>%
  pivot_longer(cols = !race_description, names_to = "type", values_to = "risk") %>% filter(race_description != "Unknown")

ggplot(stack, aes(x = race_description, y = risk)) +
  geom_bar(stat = "identity", width= 0.5, fill = "#21aca9") +
  facet_wrap(~ type, ncol = 1, scales = "free_y")+
  xlab("Race") + 
  ylab("Top 3 most frequent SMM indicator rate") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 0.2)) 
 # geom_text(aes(label = severe_21_perc),size = 3, vjust = -0.5, position = position_dodge(0.9)) +
  theme(axis.text.x = element_text(size=7,angle = 45, vjust = 1, hjust=1),
        axis.text.y = element_text(size = 8),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "grey80"),
       panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey80"),
        panel.background = element_blank())



```
```{r}
hyst

ggplot(hyst[-c(5),], aes(x = race_description, y = trans_risk)) + 
  geom_bar(stat = "identity", width= 0.5, fill = "#029076") +
  scale_fill_gradient()+
  xlab("") + 
  ylab("Transfusion Risk") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 0.2)) + 
  scale_x_discrete(labels=c("Black" = "Black**", "White" = "White(ref)", "Hispanic" =  "Hispanic**", "Asian or PI"= "Asian or PI**", "Native American" = "Native American","Other" = "Other**", "All"))+
 # geom_text(aes(label = severe_21_perc),size = 3, vjust = -0.5, position = position_dodge(0.9)) +
  theme(axis.text.x = element_text(size= 6),
        axis.text.y = element_text(size = 8),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "grey80"),
       panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey80"),
        panel.background = element_blank(), 
       aspect.ratio = 0.3)
       #panel.border = element_rect(size = 0.5, linetype = 'solid', colour = "grey80", fill = NA))


ggplot(hyst[-c(5),], aes(x = race_description, y = hyst_risk)) + 
  geom_bar(stat = "identity", width= 0.5, fill = "#029076") +
  scale_fill_gradient()+
  xlab("Race") + 
  ylab("Hysterectomy Risk") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.5), limits = c(0, 0.02)) + 
  scale_x_discrete(labels=c("Black" = "Black**", "White" = "White(ref)", "Hispanic" =  "Hispanic*", "Asian or PI"= "Asian or PI", "Native American" = "Native American","Other" = "Other", "All"))+
 # geom_text(aes(label = severe_21_perc),size = 3, vjust = -0.5, position = position_dodge(0.9)) +
  theme(axis.text.x = element_text(size= 6),
        axis.text.y = element_text(size = 8),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "grey80"),
       panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey80"),
        panel.background = element_blank(), 
       aspect.ratio = 0.3)
       #panel.border = element_rect(size = 0.5, linetype = 'solid', colour = "grey80", fill = NA))




```
Figure 3. The risk for morbidity based on the median income of patients’ zipcode.
```{r}
inc_table <- data.frame (zipinc_qrtl = c(1,2,3,4,NA),
                  inc_description = c("1", "2", "3", "4", "Unknown") )

PPH <- left_join(PPH, inc_table, by = "zipinc_qrtl" )

inc_stats <- PPH %>% group_by(inc_description) %>% summarise(count = n(), demograph_perc = round(n()/117583, digits = 3)) %>% 
  arrange(desc(demograph_perc)) %>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"All")))

SMM_inc <- PPH %>% filter(SMM == "Y") %>% group_by(inc_description) %>% summarise(smm_count = n())%>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"All")))

SMM_inc <- left_join(inc_stats, SMM_inc, by = "inc_description")
SMM_inc <- SMM_inc %>% 
  mutate(smm_rate = smm_count/count)

```


```{r}

year_pop <- PPH %>% group_by(year) %>% summarise(PPH_population_count = n())
 
year_pop$core_count <- c(7135090, 7159694, 7105498, 7083805)


year_pop <- year_pop %>% mutate(PPH_perc = PPH_population_count/core_count * 1000)

year_pop 


counts <- c('26,588', '28,375', '30,204', '32,416')
ggplot(year_pop, aes(x = year, y = PPH_perc)) +
  geom_line(color = "#5ec9ab")+
  geom_point(size = 2, color = "#267f87")+
  xlab("Year") + 
  ylab("Number of PPH patients\nper 1000 inpatient records\n") + geom_text(aes(label = counts),size = 3, vjust = 2, position = position_dodge(0.9))+
  scale_y_continuous(limits = c(3, 5)) + 
  theme_minimal()
```

```{r}
PPH %>% filter(SMM == "Y") %>% summarize(n()) 
# 21501 out of 117583 are severe
cdc_severe_count <- c(sum(PPH$cdc_s1), sum(PPH$cdc_s2),sum(PPH$cdc_s3), sum(PPH$cdc_s4),sum(PPH$cdc_s5), sum(PPH$cdc_s6),sum(PPH$cdc_s7), sum(PPH$cdc_s8),sum(PPH$cdc_s9), sum(PPH$cdc_s10),sum(PPH$cdc_s11), sum(PPH$cdc_s12),sum(PPH$cdc_s13), sum(PPH$cdc_s14),sum(PPH$cdc_s15), sum(PPH$cdc_s16),sum(PPH$cdc_s17), sum(PPH$cdc_s18),sum(PPH$cdc_s19), 
sum(PPH$cdc_s20),sum(PPH$cdc_s21))

cdc_severe_labels <- sprintf("cdc_s%d",seq(1:21))

cdc_smm <- as.data.frame(cdc_severe_count,cdc_severe_labels) %>% arrange(desc(cdc_severe_count)) %>% mutate(cdc_severe_rate = cdc_severe_count/21501) %>% head(6)
cdc_smm$name <- c("Transfusion", "Disseminated\nintravascular coagulation","Hysterectomy", "Sickle cell\ndisease with crisis","Acute renal\nfailure","Adult respiratory\ndistress syndrome")


ggplot(cdc_smm, aes(x = reorder(name, cdc_severe_rate) , y = cdc_severe_rate)) +
  geom_bar(stat = "identity", width= 0.5, fill = "#21aca9") +
  coord_flip()+
  xlab("SMM Indicators") + 
  ylab("Percentage of single SMM indicators") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + 
  theme_minimal()
```

Among the SMM patients, the most frequent CDC dx/pr are 17. Transfusion, 8. Disseminated intravascular coagulation (DIC), 19.  Hysterectomy, 15.Sickle cell disease with crisis

### drafts below
```{r}

race_percentage_plot <- ggplot(PPH, aes(race)) + 
          geom_bar(aes(y = (..count..)/sum(..count..))) + 
          scale_y_continuous(labels=scales::percent) +
  ylab("race percentage")


race_percentage_plot
```

```{r}
library(scales) 
data(tips, package = "reshape2")
myplot <- ggplot(tips, aes(day)) + 
          geom_bar(aes(y = (..count..)/sum(..count..))) + 
          scale_y_continuous(labels=scales::percent) +
  ylab("relative frequencies")

myplot

```


```{r}
PPH <- PPH %>% 
  mutate(SMM = ifelse(!!rowSums(PPH[126:146]), "Y", "N")) 

two_severe <- PPH %>% group_by(SMM, aprdrg_severity) %>% summarize(total = n()) 

PPH %>% ggplot(aes(x = aprdrg_severity)) + geom_bar() + facet_wrap(~ SMM)

PPH %>% summarize(n())

```

CDC's Severe Maternal Morbidity and aprdrg_severity

### Figure 1. The risk for severe maternal morbidity (SMM) based on race, including transfusion (left) and roldugin transfusion (right) 

```{r}
sever <- PPH %>% filter(sever_morbidity == 1) 
a <- sever %>% group_by(race) %>% summarise(s = n())
b <- PPH %>% group_by(race) %>% summarise(total = n())
c <- cbind(a, b$total) %>% mutate(percentage = s/b$total)
```

### Figure 2. The risk for disseminated intravascular coagulation (A), transfusion (B), and hysterectomy (C) within a race for patients with postpartum hemorrhage.


### Figure 3. The risk for morbidity based on the median income of patients’ zipcode.



