---
title: "replicate PPH paper"
output: pdf_document
---

```{r}
library(tidyverse)
library(tidymodels)
library(readr)
PPH <- read_csv("PPH_newcols.csv") %>% mutate(SMM = ifelse(!!rowSums(PPH[126:146]), "Y", "N"))

race_table <- data.frame (race = c(1,2,3,4,5,6,NA),
                  race_description = c("White", "Black", "Hispanic", "Asian or PI", "Native American", "Other", "Unknown") )

PPH <- left_join(PPH, race_table, by = "race" )

head(PPH)
```

"A total of 360,370 (3.2%) of 11,260,869
estimated deliveries to women aged
15e54 years complicated by postpartum
hemorrhage were included in this anal-
ysis (Table 1)."

```{r}
# We have 117583 observation of PPH patients.
# If we take weight into consideration, then we have

PPH %>% select(discwt) %>% sum()
```
PPH patient count by racce
```{r}

race_stats <- PPH %>% group_by(race_description) %>% summarise(count = n(), demograph_perc = round(n()/117583, digits = 3)) %>% arrange(desc(demograph_perc))

#hide the row for unknown
race_stats_known <- race_stats[-c(5),]


```

```{r}
library(scales)
ggplot(race_stats_known, aes(x = reorder(race_description, -demograph_perc), y = demograph_perc)) +
  geom_bar(stat = "identity", width= 0.8, fill = "blue") +
  xlab("Race") + 
  ylab("Percentage ") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8),)

```

### PPH Mortality by Race

```{r}
died <- PPH %>% filter(died == 1) %>% group_by(race_description) %>% summarise(mortality = n()) 

died <- left_join(race_stats, died, by = "race_description")
died <- died %>% 
  mutate(mort_rate = mortality/race_stats$count)


ggplot(died[-c(5),], aes(x = reorder(race_description, -mort_rate), y = mort_rate)) +
  geom_bar(stat = "identity", width= 0.8, fill = "blue") +
  xlab("Race") + 
  ylab("Mortality Rate ") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(axis.text.x = element_text(size = 7), 
        axis.text.y = element_text(size = 8))

```

```{r}

race_percentage_plot <- ggplot(PPH, aes(race)) + 
          geom_bar(aes(y = (..count..)/sum(..count..))) + 
          scale_y_continuous(labels=scales::percent) +
  ylab("race percentage")


race_percentage_plot
```

```{r}
library(scales) 
data(tips, package = "reshape2")
myplot <- ggplot(tips, aes(day)) + 
          geom_bar(aes(y = (..count..)/sum(..count..))) + 
          scale_y_continuous(labels=scales::percent) +
  ylab("relative frequencies")

myplot

```


```{r}
PPH <- PPH %>% 
  mutate(SMM = ifelse(!!rowSums(PPH[126:146]), "Y", "N")) 

two_severe <- PPH %>% group_by(SMM, aprdrg_severity) %>% summarize(total = n()) 

PPH %>% ggplot(aes(x = aprdrg_severity)) + geom_bar() + facet_wrap(~ SMM)

PPH %>% summarize(n())

```

CDC's Severe Maternal Morbidity and aprdrg_severity

### Figure 1. The risk for severe maternal morbidity (SMM) based on race, including transfusion (left) and roldugin transfusion (right) 

```{r}
sever <- PPH %>% filter(sever_morbidity == 1) 
a <- sever %>% group_by(race) %>% summarise(s = n())
b <- PPH %>% group_by(race) %>% summarise(total = n())
c <- cbind(a, b$total) %>% mutate(percentage = s/b$total)
```

### Figure 2. The risk for disseminated intravascular coagulation (A), transfusion (B), and hysterectomy (C) within a race for patients with postpartum hemorrhage.


### Figure 3. The risk for morbidity based on the median income of patients’ zipcode.


```{r}
x <- tibble(a = c("White", "Black", "Asian"), b = c(45, 21, 10))
x

x %>$

```
