---
title: "Untitled"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readr)

#import all data

NIS_2016_Core <- read_csv("~/datathon/NIS_2016_Core.csv") %>%
  mutate(key = round(key_nis))
NIS_2016_Severity <- read_csv("~/datathon/NIS_2016_Severity.csv") %>%
  mutate(key = round(key_nis))

NIS_2017_Core <- read_csv("~/datathon/NIS_2017_Core.csv") %>%
  mutate(key = round(key_nis))
NIS_2017_Severity <- read_csv("~/datathon/NIS_2017_Severity.csv") %>%
  mutate(key = round(key_nis))

NIS_2018_Core <- read_csv("~/datathon/NIS_2018_Core.csv") %>%
  mutate(key = round(key_nis))
NIS_2018_Severity <- read_csv("~/datathon/NIS_2018_Severity.csv") %>%
  mutate(key = round(key_nis))

NIS_2019_Core <- read_csv("~/datathon/NIS_2019_Core.csv") %>%
  mutate(key = round(key_nis))
NIS_2019_Severity <- read_csv("~/datathon/NIS_2019_Severity.csv") %>%
  mutate(key = round(key_nis))


#filter out PPH patients (by O720/1/2/3 in diagnosis)

icd_2016 <- sprintf("i10_dx%d",seq(1:30)) #2016 has 10_dx1-30, the other three year has 1-40
PPH_Core_2016 <- NIS_2016_Core %>% 
  filter_at(icd_2016, any_vars(.%in% c("O720","O721", "O722", "O723"))) #26588 patients diagonsed with PPH in 2016

icd_2017 <- sprintf("i10_dx%d",seq(1:40)) 
PPH_Core_2017 <- NIS_2017_Core %>% 
  filter_at(icd_2017, any_vars(.%in% c("O720","O721", "O722", "O723"))) #28375 patients diagonsed with PPH in 2017

icd_2018 <- sprintf("i10_dx%d",seq(1:40))
PPH_Core_2018 <- NIS_2018_Core %>% 
  filter_at(icd_2018, any_vars(.%in% c("O720","O721", "O722", "O723"))) #30204 patients diagonsed with PPH in 2018

icd_2019 <- sprintf("i10_dx%d",seq(1:40)) 
PPH_Core_2019 <- NIS_2019_Core %>% 
  filter_at(icd_2019, any_vars(.%in% c("O720","O721", "O722", "O723"))) #32416 patients diagonsed with PPH in 2019

#Join severity table, combine 2016-2019

PPH_2016 <- left_join(PPH_Core_2016, NIS_2016_Severity, by = "key") %>%     select(-c("key_nis.x","hosp_nis.y")) %>% 
  select(-c("dxver","i10_ecause1","i10_ecause2", "i10_ecause3" ,"i10_ecause4" ,"i10_necause","prver")) %>% 
  add_column("i10_dx31" = NA,"i10_dx32" = NA,"i10_dx33" = NA,"i10_dx34"= NA, "i10_dx35"= NA,"i10_dx36"= NA,"i10_dx37"= NA, "i10_dx38"= NA,"i10_dx39"= NA,"i10_dx40"= NA, .after = "i10_dx30") %>%
  add_column("i10_pr16" = NA, "i10_pr17" = NA,"i10_pr18" = NA,"i10_pr19"= NA,"i10_pr20"= NA,"i10_pr21"= NA,"i10_pr22"= NA,"i10_pr23"= NA, "i10_pr24"= NA,"i10_pr25"= NA, .after = "i10_pr15") %>%
  add_column("prday16"= NA , "prday17"= NA,  "prday18"= NA , "prday19" = NA,
"prday20" = NA,"prday21"= NA, "prday22"= NA,"prday23"= NA, "prday24"= NA, "prday25"= NA, .after = "prday15")
#the variables that 2016 has but no longer exists for later years："dxver""i10_ecause1" "i10_ecause2" "i10_ecause3" "i10_ecause4" "i10_necause" "prver"， remove them for now

PPH_2017 <- left_join(PPH_Core_2017, NIS_2017_Severity, by = "key") %>% select(-c("key_nis.x","hosp_nis.y","dxver","prver")) #the dxver and prver in 2017 are all 10,no these two columns in years after

PPH_2018 <- left_join(PPH_Core_2018, NIS_2018_Severity, by = "key") %>% select(-c("key_nis.x","hosp_nis.y"))

PPH_2019 <- left_join(PPH_Core_2019, NIS_2019_Severity, by = "key") %>% select(-c("key_nis.x","hosp_nis.y")) %>% 
select(-c("i10_birth", "i10_delivery", "i10_injury",    "i10_multinjury" , "i10_serviceline" ,"pclass_orproc")) #there are 6 new variables in 2019 "i10_birth", "i10_delivery", "i10_injury"     "i10_multinjury" , "i10_serviceline" ,"pclass_orproc", drop them for now


#final,big dataset combining 2016-2019 year
PPH <- rbind(PPH_2016, PPH_2017, PPH_2018, PPH_2019)

```

### Visualizations
```{r}
library(readr)
library(tidyverse)
PPH <- read_csv("PPH.csv")

# race stats
race_dist <- PPH %>% group_by(race) %>% summarise(count = n(), percentage = n()/117583) %>% add_column("description"= c("White", "Black", "Hispanic", "Asian or Pacific Islander", "Native American", "Other", "Missing"), .after = "race")


race_dist %>% 
ggplot(aes(x=description, y=percentage)) + 
  geom_bar(stat = "identity")

#https://www.r-graph-gallery.com/218-basic-barplots-with-ggplot2.html
```

### some tables
```{r}
# severity level distribution
sev_lev <- PPH %>% group_by(aprdrg_severity) %>% summarize(n())
sev_lev

# Gestational hypertension
gest_hyp <- PPH %>% 
  filter_at(i10_dxs, any_vars(.%in% c("O13", "O131", "O133", "O139"))) 

```

```{r}
#How to generate a binary variable based on occurence of a patter in each row

# stackoverflow  https://stackoverflow.com/questions/47941680/grepl-across-multiple-specified-columns
AB <- c('CHINAS PARTY CONGRESS','JAPAN-US RELATIONS','JAPAN TRIES TO')
TI <- c('AMERICAN FOREIGN POLICY', 'CHINESE ATTEMPTS TO', 'BRITAIN HAS TEA')
AU <- c('AUTHOR 1', 'AUTHOR 2','AUTHOR 3')
M  <- data.frame(AB,TI,AU)
M$China <- !!rowSums(sapply(M[1:2], grepl, pattern = "CHINA|CHINESE|SINO"))


#mini dx trial
t <- c("O69", "O72005", "O721","O72", "O7221","O72211", "O723")
y <- sapply(t, grepl, pattern = "^O72") #look for everything that begins with O72
z <- sapply(t, grepl, pattern = "O723|^O69") #look for the ones that are o723 or starts with o69

u <- c("O527", "O728", "O778","O72091", "O456","O3872", "O72096")

v <- c("O69", "E668", "E67","O720", "O72201","O72213", "O725")

df <- rbind.data.frame(t,u,v)
colnames(df) <- sprintf("i10_dx%d",seq(1:7))

df

df$A <- ifelse(!!rowSums(sapply(df[1:7], grepl, pattern = "E66")), 1,0) #pattern means contain, not exact, so E668 still returns true

#to find exact match, use start^ and end$

df$B <- ifelse(!!rowSums(sapply(df[1:7], grepl, pattern = "^O722$")), 1,0) #see if any cell in a row contains exactly o722, which has none
df$C <- ifelse(!!rowSums(sapply(df[1:7], grepl, pattern = "O722")), 1,0) #without noting start and end, it searches for any string that has O722 in it.
df$D <- ifelse(!!rowSums(sapply(df[2:7], grepl, pattern = "^O72005$|^E6")), 1,0) #multiple conditions

```

### Create variables for other predictors

1. Obesity: O99.214; Unspecified obesity E66.9

```{r}

i10_dxs <- sprintf("i10_dx%d",seq(1:40))

PPH$obesity <- ifelse(!!rowSums(sapply(PPH[17:56], grepl, pattern = "^O99214$|^E669$")),1,0)

#quick check
o <- PPH %>% select(i10_dxs, obesity) %>% filter(obesity == 1) #15854 obese
o2 <- PPH %>% 
  filter_at(i10_dxs, any_vars(.%in% c("O99214","E669"))) 
#using the other method, got the same dataset
```



### Create variables for Severe Maternal Morbidity (CDC)

1. Acute myocardial infarctionTransfusion: I21.01, I21.02, I21.09, I21.11, I21.19, I21.21, I21.29, I21.3, I21.4, I21.9, I21.A1 and I21.A9(I21.xx, I22.x)
 
```{r}
PPH$AMI <- ifelse(!!rowSums(sapply(PPH[17:56], grepl, pattern = "^I21|^I22")),1,0)
PPH$AMI2 <- ifelse(!!rowSums(sapply(PPH[17:56], grepl, pattern = "I2101|I2102|I2109|I2111|I2119|I2121|I2129|I213|I214|I219|I21A1|I21A9|I220|I221|I222|I228|I229")),1,0) #no spaces in between allowed

#quick check
a1 <- PPH %>% select(i10_dxs, AMI) %>% filter(AMI == 1) #26 ami
a2 <- PPH %>% select(i10_dxs, AMI2) %>% filter(AMI2 == 1) #26 ami, the same

PPH <- PPH %>% select(-AMI2)
```

2.  Aneurysm: I71.xx,I79.0
```{r}
PPH$Ane <- ifelse(!!rowSums(sapply(PPH[17:56], grepl, pattern = "^I71|^I790$")),1,0)

PPH %>% select(i10_dxs, Ane) %>% filter(Ane == 1) #8 Ane
```

3. Acute renal failure: N17.0, N17.1, N17.2, N17.8, N17.9, O90.4 (N17.x, O90.4)
```{r}
PPH$ARF <- ifelse(!!rowSums(sapply(PPH[17:56], grepl, pattern = "^N17|^O904$")),1,0)

PPH %>% select(i10_dxs, ARF) %>% filter(ARF == 1) #1099 NRF
```

4. Adult respiratory distress syndrome: J80, J95.1, J95.2, J95.3, J95.821, J95.822, J96.00, J96.01, J96.02,
J96.20, J96.21, J96.22, R09.2 (J80, J95.1, J95.2, J95.3, J95.82x, J96.0x, J96.2x R09.2)
```{r}
PPH$ARDS <- ifelse(!!rowSums(sapply(PPH[17:56], grepl, pattern = "^J80$|^J951$|^J952$|^J953$|^J9582|^J960|^J962|^R092$")),1,0)

```

5. Amniotic fluid embolism: O88.111,O88.112,O88.113, O88.12 (childbirth), O88.13

6. Cardiac arrest/ventricular fibrillation: I46.2, I46.8, I46.9, I49.01, I49.02

7. Conversion of cardiac rhythm(PR):	5A2204Z, 5A12012

8. Disseminated intravascular coagulation: D65, D68.8, D68.9, O72.3

9. Eclampsia:	O15.00, O15.02, O15.03, O15.1, O15.2, O15.9

10. Heart failure/arrest during surgery or procedure: I97.120, I97.121, I97.130, I97.131, I97.710, I97.711

12. Pulmonary edema / Acute heart failure: J81.0, I50.1, I50.20, I50.21, I50.23, I50.30, I50.31, I50.33, I50.40, I50.41, I50.43, I50.9

13. Severe anesthesia complications: O74.0 , O74.1, O74.2, O74.3, O89.01, O89.09, O89.1, O89.2

14. 

16. Sickle cell disease with crisis: D57.00 , D57.01, D57.02, D57.211, D57.212, D57.219, D57.411, D57.412, D57.419, D57.811, D57.812, D57.819

17. Air and thrombotic embolism: I26.01, I26.02, I26.09, I26.90, I26.92, I26.99, O88.011-O88.019, 088.02, O88.03, O88.211-O88.219, O88.22, O88.23, O88.311-O88.319, O88.32, O88.33, O88.81, O88.82, O88.83

18. Blood products transfusion(PR): 

19. Hysterectomy(PR): 0UT90ZZ, 0UT94ZZ, 0UT97ZZ, 0UT98ZZ, 0UT9FZZ

20. Temporary tracheostomy(PR): 0B110Z4, 0B110F4, 0B113Z4, 0B113F4, 0B114Z4, 0B114F4

21.Ventilation(PR): 5A1935Z, 5A1945Z,
5A1955Z



```{r}

PPH$test <- grepl("O", PPH$i10_dx1), 1, 0)
i10_dxs <- sprintf("i10_dx%d",seq(1:40))
i10_prs <- sprintf("i10_pr%d",seq(1:25))
A <- PPH %>% select(i10_prs)

A$trans <- sapply(A, str_starts, "30233|10Z")
B <- A %>% filter(trans == TRUE)

A$t <- str_starts(A, "30233")
#PPH %>% select(i10_dx1, test)


# create an additional column indicating number of times transfusion is taken among all procedures
PPH$transfusion <- ifelse(rowSums(sapply(PPH[59:83], grepl, pattern = "30233H1|30233L1|30233K1|30233M1|30233N1|30233P1|30233R1|30233T1|30233H0|30233L0|30233K0|30233M0|30233N0|30233P0|30233R0|30233T0")) == 0, 0,1)
PPH$t2 <- ifelse(rowSums(sapply(PPH[59:83], str_starts, pattern = "30233H1|30233L1|30233K1|30233M1|30233N1|30233P1|30233R1|30233T1|30233H0|30233L0|30233K0|30233M0|30233N0|30233P0|30233R0|30233T0")) == 0, 0,1)

transfu <- PPH %>% filter(transfusion == 1)

t2<- PPH %>% filter(t2 == 1)
# create an additional column indicating obesity in diagonsis
PPH$obesity <- ifelse(sapply(PPH[17:56], grepl, pattern = "O99214"),1,0)

#quick check
a <- PPH %>% select(obesity, sprintf("i10_dx%d",seq(1:40))) %>% filter(obesity == 1)

# Multiple gestation: O30.8, O30.8x, O30.9
PPH$MG <- ifelse(sapply(PPH[17:56], grepl, pattern = "O30"),1,0)
PPH %>% select()


n <- PPH %>% filter(drg != drg_nopoa)

```